A numerical study of the solution of X-Mode equations around the hybrid resonance

### 2015-02-10 report_clean/numerics.tex
numerics.tex: removed Condition number paragraph

We use the following ingredients:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{h,\nu}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}
 \leq C\nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V},\; C>0.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness} and is valid for $\nu\rightarrow 0$.   
\item The form of the exact solution to the problem (\ref{vf1dcase}), see \cite{Despres_2014},
\begin{align}
\label{eq:exact}
 E_x^{\nu}=-iE_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f^{\nu}(x)}{\alpha(x)+i\nu}, \; \text{for some }f^{\nu}(x)\in L_{2}(\Omega).
\end{align}
\mrev{We additionally assume that $E_y^{\nu}(x)$ is sufficiently smooth and $\|(E_y^{\nu})''\|_{H_{2}(\Omega)}$
can be bounded independently of $\nu$}.
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:interp}
\begin{split}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2\left|v''\right|_{L_{2}(\Omega)},\; C>0,\\
 \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
 \end{split}
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$. 
\end{itemize}
\mrev{The estimate (\ref{eq:cea}) combined with (\ref{eq:interp}) results in, for some $C>0$:
\begin{align}
  \label{eq:final_est}
  \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}%\leq C_1\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L_{2}}+C_2\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}}
  \leq C\nu^{-1}h+C\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}},
\end{align}
where we used an assumption that the $L_2$-norm of $\left(E_y^{\nu}(x)\right)''$ can be bounded independently of $\nu$.} For $E_x^{\nu},\; f^{\nu}(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\ben
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{(f^{\nu})''}{\alpha+i\nu}-2\frac{(f^{\nu})'\alpha'}{(\alpha+i\nu)^2}+\frac{f^{\nu}\alpha''}{(\alpha+i\nu)^3},
\een
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\bealn
 \left|\frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
 % &=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\eealn
where $C>0$ does not depend on $\nu$. After inserting this into (\ref{eq:final_est}) we obtain
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2+C\nu^{-1}h.
\een
From this it thus follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. 
This prediction is very severe and is due to the singular nature of the problem under consideration.

@@

\begin{figure}[htb!]
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/fig_cond_num.pdf}&
 \includegraphics[height=0.32\textwidth]{pics_frequency_domain/ey_resonance.pdf}
 \end{tabular}
 \caption{
 The left plot demonstrates the dependence of the condition number of the system (\ref{eq:simple_system}) on $\nu$, for different values of $h$ (the condition 
 number for $\nu=0$ is not shown, however, for all values of $h$ as shown in the plot, the matrix was non-singular even for $\nu=0$). 
 The right plot shows $E_{y}^{\nu}$ computed on different meshes for small $\nu$. It can be seen that for meshes that are coarse 
 with respect to $\nu$ (e.g. $\nu=10^{-4}, \; h=10^{-2}$, or $\nu=0,\; h=10^{-2}$), the computed solution seems to be only piecewise-continuous in $\nu$, 
 while for finer meshes and $\nu>0$ the solution seems to be smooth. It can be also seen that for $h=10^{-2}$ and $h=10^{-3}$ the solutions 
 computed for $\nu=0$ are almost indistinguishable.}
 \label{fig:small_nu}
\end{figure}
%\begin{table}[ht!]
%\begin{tabular}{c|ccccccccc}
%\diagbox[width=10em]{$h$}{$\nu$}& 100 & 10 & 1 & 0.1 & 1e-2 & 1e-3 & 1e-4 & 1e-8 & 0\\
% \hline
%1 & 3.9634&4.3787 &4.987&36.949&42.394& 42.757 &42.79&42.793&42.793\\
%\hline
%0.5&      3.9931& 6.87&48.473 & 168.41  &207.34  & 209.63 & 209.83 & 209.85&  209.85\\
%\hline
% 0.1&      16.09  &     159.56  &     1146.1   &    8304.1   &     19169  &      20270   &     20344   &     20352   &     20352\\
% \hline
%0.05&      64.046  &     637.18  &     4686.2   &     38876 &  1.4e5 &  1.58e5 &   1.58e5 &  1.58e5 &  1.58e5\\
%\hline
% 0.01&       1600    &   15921 & 1.2e5  & 1.1e6 & 8.3e6  & 1.8e7   & 1.93e7  & 1.93e7  & 1.93e7\\
%\end{tabular}
%\caption{The condition number of the matrix of the system (\ref{eq:simple_system}) for different values of $\nu$ and $h$. }
%\label{tab:cond_number}
%\end{table}
\paragraph{Condition Number}
In Figure \ref{fig:small_nu} we demonstrate the dependence of the condition number of the matrix of the system (\ref{eq:simple_system}) 
on $\nu$, for several values of $h$.
Remarkably, for $\nu=0$ the computed matrices are not singular. We do not know the exact reason for this.
As an additional illustration, we compare the solutions for very small $\nu$ and $\nu=0$ for different meshes in Fig.~\ref{fig:small_nu}.

>>>>

We use the following ingredients:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{h,\nu}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}
 \leq C\nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V},\; C>0.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness} and is valid for $\nu\rightarrow 0$.   
\item The form of the exact solution to the problem (\ref{vf1dcase}), see \cite{Despres_2014},
\begin{align}
\label{eq:exact}
 E_x^{\nu}=-iE_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f^{\nu}(x)}{\alpha(x)+i\nu}, \; \text{for some }f^{\nu}(x)\in L_{2}(\Omega).
\end{align}
We additionally assume that $E_y^{\nu}(x)$ is sufficiently smooth and $\|(E_y^{\nu})''\|_{H_{2}(\Omega)}$
can be bounded independently of $\nu$. 
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:interp}
\begin{split}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2\left|v''\right|_{L_{2}(\Omega)},\; C>0,\\
 \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
 \end{split}
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$. 
\end{itemize}
The estimate (\ref{eq:cea}) combined with (\ref{eq:interp}) results in, for some $C>0$,
\begin{align}
  \label{eq:final_est}
  \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}%\leq C_1\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L_{2}}+C_2\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}}
  \leq C\nu^{-1}h+C\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}},
\end{align}
where we used an assumption that the $L_2$-norm of $\left(E_y^{\nu}(x)\right)''$ can be bounded independently of $\nu$.

For $E_x^{\nu},\; f^{\nu}(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\ben
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{(f^{\nu})''}{\alpha+i\nu}-2\frac{(f^{\nu})'\alpha'}{(\alpha+i\nu)^2}+\frac{f^{\nu}\alpha''}{(\alpha+i\nu)^3},
\een
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\bealn
 \left|\frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
 % &=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\eealn
where $C>0$ does not depend on $\nu$. After inserting this into (\ref{eq:final_est}) we obtain
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2+C\nu^{-1}h.
\een
From this it thus follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. 
This prediction is very severe and is due to the singular nature of the problem under consideration.

@@

In Figure \ref{fig:small_nu} we demonstrate the dependence of the condition number of the matrix of the system (\ref{eq:simple_system}) 
on $\nu$, for several values of $h$.
Remarkably, for $\nu=0$ the computed matrices are not singular. We do not know the exact reason for this.
As an additional illustration, we compare the solutions for very small $\nu$ and $\nu=0$ for different meshes in Fig.~\ref{fig:small_nu}. 
\begin{figure}[htb!]
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/fig_cond_num.pdf}&
 \includegraphics[height=0.32\textwidth]{pics_frequency_domain/ey_resonance.pdf}
 \end{tabular}
 \caption{
 The left plot demonstrates the dependence of the condition number of the system (\ref{eq:simple_system}) on $\nu$, for different values of $h$ (the condition 
 number for $\nu=0$ is not shown, however, for all values of $h$ as shown in the plot, the matrix was non-singular even for $\nu=0$). 
 The right plot shows $E_{y}^{\nu}$ computed on different meshes for small $\nu$. It can be seen that for $h=10^{-2}$ and $h=10^{-3}$ the solutions 
 computed for $\nu=0$ are almost indistinguishable.}
 %It can be seen that for meshes that are coarse 
 %with respect to $\nu$ (e.g. $\nu=10^{-4}, \; h=10^{-2}$, or $\nu=0,\; h=10^{-2}$), the computed solution seems to be only piecewise-continuous in $\nu$, 
 %while for finer meshes and $\nu>0$ the solution seems to be smooth. }
 \label{fig:small_nu}
\end{figure}
%\begin{table}[ht!]
%\begin{tabular}{c|ccccccccc}
%\diagbox[width=10em]{$h$}{$\nu$}& 100 & 10 & 1 & 0.1 & 1e-2 & 1e-3 & 1e-4 & 1e-8 & 0\\
% \hline
%1 & 3.9634&4.3787 &4.987&36.949&42.394& 42.757 &42.79&42.793&42.793\\
%\hline
%0.5&      3.9931& 6.87&48.473 & 168.41  &207.34  & 209.63 & 209.83 & 209.85&  209.85\\
%\hline
% 0.1&      16.09  &     159.56  &     1146.1   &    8304.1   &     19169  &      20270   &     20344   &     20352   &     20352\\
% \hline
%0.05&      64.046  &     637.18  &     4686.2   &     38876 &  1.4e5 &  1.58e5 &   1.58e5 &  1.58e5 &  1.58e5\\
%\hline
% 0.01&       1600    &   15921 & 1.2e5  & 1.1e6 & 8.3e6  & 1.8e7   & 1.93e7  & 1.93e7  & 1.93e7\\
%\end{tabular}
%\caption{The condition number of the matrix of the system (\ref{eq:simple_system}) for different values of $\nu$ and $h$. }
%\label{tab:cond_number}
%\end{table}
